\begin{abstract}
Safe RL systems usually fail when regime shifts are faster than policy adaptation. In governance-like control loops, that failure does not first appear as lower average reward; it appears as brittle decisions under uncertainty.

We study an uncertainty-gated safe RL controller that routes PPO actions through a calibrated uncertainty gate and a deterministic shield. The core contribution is a control architecture claim: uncertainty-gated routing plus deterministic shielding gives a better safety-stability operating point under shift than policy-only control, while a compact classically simulated VQC-style head serves as a parameter-efficient non-linear uncertainty estimator.

Our primary empirical evidence comes from standard constrained benchmarks (SafetyPoint\-Goal1-v0 and SafetyHalf\-Cheetah\-Velocity-v1; 30 seeds) plus 50-seed governance controller comparisons under matched protocols. Custom governance simulations are treated as stress tests, not as the sole source of external validity.

We also report a negative result directly: in one custom ablation workload, full-system reward (-1264.90) is below random (-1232.33). We keep this result because it clarifies what the method is and is not optimizing: the method improves control stability and safety-oriented behavior under shift, but does not guarantee reward dominance in every stylized reward design. The central claim is therefore parameter efficiency with comparable safety behavior rather than universal return superiority. In a separate larger-model efficiency protocol, the compact architecture achieves a 166.89$\times$ parameter reduction while remaining competitive on safety-oriented metrics.

Code and artifacts: \url{https://github.com/maverick0721/QAI-Chain.git}.
\end{abstract}
