\section{Discussion}

The headline interpretation is not ``we found a universally better learner''; it is that a small decision-time safety layer can change what failure looks like under shift. Read the plots as trade-offs: where a policy-only controller is brittle, gating plus a deterministic shield can trade some raw return for predictability and constraint adherence.

External validity should be judged first on standard constrained tasks. SafetyPointGoal1-v0 and SafetyHalfCheetahVelocity-v1 under a fixed 30-seed protocol establish that the training and evaluation loop behaves sanely outside the custom governance simulator. The governance environments then serve as stress tests designed to surface regime-switch brittleness and miscalibration.

The negative result is also part of the message. In one stylized ablation workload, random reward is less negative than the full system. We keep that table because it is a concrete reminder that reward design can be gamed and that ``reward rank'' is not a sufficient certificate in safety-critical control. The gated controller often prefers conservative, shield-compatible actions; in some reward designs that can look worse on the scalar return while still behaving better on tracking and constraint-oriented metrics.

Where the method helps is when switching intensity rises. Under those conditions, routing by calibrated uncertainty improves the precision of ``when to fall back'' without requiring large ensembles, and the deterministic shield limits tail-risk trajectories even when the gate makes mistakes. This is why the efficiency mechanism matters: a compact uncertainty head can be used as a practical signal inside the loop rather than an offline diagnostic.

Positioning-wise, this is a systems-and-control pattern that can sit on top of modern safe-RL objectives; it does not replace them. The OmniSafe constrained baselines reinforce the broader lesson: return rankings move with budgets and regimes, while safety costs often behave like a frontier, not a leaderboard.

\subsection{Why Uncertainty-Gated Control Works}

The mechanism can be understood as three interacting components.

First, uncertainty gating improves branch selection under shift. When epistemic uncertainty rises, routing to a deterministic heuristic avoids committing to poorly calibrated policy logits in exactly the regime where policy-only execution is most brittle.

Second, shielding regularizes action execution even when branch selection is imperfect. Bounded actions and bounded action-rate change constrain tail-risk trajectories that otherwise dominate safety incidents in high-switch regimes.

Third, risk-shaped training complements decision-time controls. Penalizing short-window reward variance reduces incentive to exploit transient reward spikes that are not aligned with stable control intent.

This decomposition clarifies why gains are most visible under disturbance and less pronounced in easier stationary regimes: the architecture contributes primarily through failure-mode control, not through unconditional reward amplification.

\subsection{Failure Cases}

The controller still fails under two regimes. First, when uncertainty calibration drifts faster than the gate can track, fallback can trigger late and violations can spike briefly before stabilizing. Second, when disturbances exceed the shield's modeled envelope, the policy and heuristic can oscillate, degrading tracking.

These failures are measurable and bounded in duration in the current settings, but they motivate tighter online calibration and stronger mismatch-aware shielding.
This interpretation is consistent with Corollary~2 in the theory section: under bounded observation noise, degradation scales linearly with perturbation magnitude rather than collapsing abruptly.

Limitations remain clear. Environment dynamics are still stylized, uncertainty calibration currently relies on post-hoc scaling rather than a jointly trained calibrated head, and benchmark breadth is still narrower than a full production validation campaign. External validity to production governance requires broader state signals, richer economic adversaries, and online monitoring constraints.

Near-term follow-up should focus on three directions:
\begin{itemize}
	\item calibrated uncertainty estimators (for example, ensemble or Bayesian approximations) instead of raw action-variance proxies,
	\item explicit safety constraints in addition to fallback routing,
	\item and transfer tests where policies trained in one disturbance profile are evaluated in another without retuning.
\end{itemize}

For deployment-minded readers, the practical message is straightforward: keep a deterministic fallback in the control loop, gate learned actions by uncertainty, and report safety-oriented metrics next to reward \citep{pineau2021improving,dror2018hitchhikers}.

\begin{table}[htbp]
\centering
\input{tables/when_to_use}
\caption{When to use this method in practice.}
\label{tab:when-to-use}
\end{table}