\section{Experimental Setup}

\paragraph{Protocol.}
We evaluate in two tiers:
\begin{itemize}
    \item \textbf{Primary constrained benchmarks:} SafetyPointGoal1-v0 and SafetyHalfCheetahVelocity-v1 under a fixed 30-seed protocol (CPO, P3O, PPOLag) with shared uncertainty-gating and shielding logic.
    \item \textbf{Governance stress-test environments:} a scaled governance simulator (12-dimensional state, 3-dimensional action, 200-step horizon, partial observability, adversarial telemetry faults) and a DeFi governance simulator (8-dimensional state, 2-dimensional action, 150-step horizon with MEV pressure and liquidity-risk dynamics).
\end{itemize}
For publication-facing comparisons, we use fixed seed sets to preserve paired comparability under stochastic dynamics.
Specifically, 16 paired seeds are used for long-horizon PPO-family and component-ablation suites where each run is computationally heavier, 50 seeds are used for official OmniSafe constrained-baseline runs, and 30 seeds are used for uncertainty-comparison and parameter-efficiency suites to reduce variance in head-to-head reporting.

\paragraph{Constrained baseline fidelity statement.}
The constrained baselines reported in the main table are official OmniSafe implementations (CPO, P3O, PPOLag, SACLag, FOCOPS) executed through OmniSafe-compatible adapters for both environments.

\paragraph{Environment realism extension.}
Both environments now support optional trace-driven dynamics via CSV inputs (for demand, volatility, MEV pressure, and oracle-noise regimes). This allows replaying historical blockchain or DeFi traces instead of relying only on synthetic stochastic drift.
For the reported real-trace run, we use Ethereum daily gas-price history from Etherscan to construct trace features that drive intensity and disturbance schedules.

\paragraph{Primary Metrics.}
We report:
\begin{itemize}
    \item mean reward and standard deviation across seeds,
    \item safety violation count,
    \item target-tracking error (MAE),
    \item fallback activation rate,
    \item per-decision latency and forward-FLOPs proxy.
\end{itemize}

Fallback activation is measured as the fraction of states where uncertainty exceeds threshold and the policy is routed to the deterministic controller before shield validation.

\paragraph{Hypothesis Testing.}
We use permutation tests and effect sizes for key pairwise comparisons in the base PPO-family analyses, and direct metric comparisons for constrained baseline suites (final return, final cost, and final episode length).

Bootstrap resampling (95\% confidence) is used for interval reporting where applicable.

\paragraph{High-rigor constrained-baseline protocol.}
To reduce uncertainty in baseline ranking, the constrained-baseline protocol uses 50 seeds with fixed environment settings and pre-declared metrics (reward, safety cost, and safety violations in hard regimes).

\paragraph{Fair-capacity quantum efficiency protocol.}
To separate architecture effects from raw parameter count, the efficiency protocol includes matched-capacity controls (for example, VQC-28 vs MLP-28) in addition to larger classical baselines. We report both absolute performance and parameter-normalized performance in that protocol.

\paragraph{Primary standard-benchmark protocol.}
The primary external-validity test uses constrained standard benchmarks (Safety Gym-style and MuJoCo-style tasks) under the same uncertainty-gating and shielding logic.
On this host, we report completed 30-seed runs on SafetyPointGoal1-v0 and SafetyHalfCheetahVelocity-v1 (CPO/P3O/PPOLag, 3k steps per run) under the OmniSafe stack.
Governance simulations are then used as controlled stress tests to probe mechanism behavior under adversarial regime shifts that are hard to isolate in standard suites.

\paragraph{Robust PPO settings.}
Robust PPO uses risk-aware shaping, quantum-uncertainty-gated fallback, and a smart-contract-style shield. The uncertainty estimator uses $K=10$ parameter-perturbed forward passes with Gaussian perturbation scale 0.01. The gate threshold $\tau_q$ and risk weight $\lambda_{\mathrm{risk}}$ are tuned for high-switch regimes.

\paragraph{Robust ablation variants.}
To isolate mechanism contributions, we evaluate seven configurations:
\begin{itemize}
    \item Full system (quantum gate + shield + robust policy),
    \item No quantum gate,
    \item No shield,
    \item No fallback,
    \item Classical PPO,
    \item Heuristic controller,
    \item Random controller.
\end{itemize}
