\section{Limitations}

This study has several important limitations:
\begin{itemize}
    \item The governance simulators are intentionally simplified and do not yet model full peer-network dynamics, endogenous market adaptation, or realistic transaction-arrival microstructure.
    \item The quantum-inspired uncertainty module and audit pipeline are research prototypes, not production-hardened systems.
    \item Runtime measurements are single-machine snapshots rather than cross-hardware scaling studies.
\end{itemize}

\par
Additional validity risks include:
\begin{itemize}
    \item reward misspecification sensitivity under alternate throughput/latency trade-offs (highlighted by the random-vs-full anomaly in one ablation workload),
    \item moderate sample size in some per-method seed analyses outside the 50-seed constrained-baseline suite,
    \item simulator realism limits for both uncertainty dynamics and network behavior,
    \item absence of richer economic-adversary models in mining and transaction dynamics.
    \item baseline coverage that is still narrower than a full 2022--2025 safe RL comparison set (for example, newer uncertainty-aware and risk-sensitive variants).
\end{itemize}

The observation-noise robustness statement (Corollary~2) is therefore a bounded-perturbation guarantee under explicit modeling assumptions, not a blanket safety certificate under arbitrary sensing failures.

The standard constrained benchmarks improve external validity, but they do not by themselves establish readiness for live governance deployment. These limitations define immediate future work rather than hidden assumptions, and they bound the claim scope to robust mechanism behavior under the tested protocols.