\section{Reproducibility Checklist}

\begin{itemize}
    \item \textbf{Code availability:} all experiment and table-generation scripts are included in the repository.
    \item \textbf{Determinism controls:} fixed seed sets are used for publication-scale runs.
    \item \textbf{Artifact generation:} JSON and markdown artifacts are generated directly from executable scripts.
    \item \textbf{Table provenance:} LaTeX tables are generated from artifact files via a dedicated script.
    \item \textbf{Validation gates:} linting, type checks, tests, and runtime health checks are automated.
\end{itemize}

The end-to-end workflow is:
\begingroup
\scriptsize
\begin{verbatim}
python experiments/run_publication_suite.py
python experiments/run_detailed_suite.py
python experiments/run_adversarial_suite.py
python experiments/run_complexity_\
regime_suite.py
python scripts/analyze_research_results.py
python scripts/generate_benchmark_report.py
python scripts/generate_paper_tables.py
python scripts/generate_appendix_tables.py
python scripts/generate_paper_figures.py
python scripts/generate_adversarial_\
artifacts.py
python scripts/generate_adversarial_\
stress_sweep.py
python scripts/generate_complexity_\
artifacts.py
\end{verbatim}
\endgroup

\paragraph{Provenance policy.}
All quantitative manuscript assets are generated from stored artifacts; manual number entry is minimized. This improves traceability and supports exact regeneration when code or protocol settings change.

\paragraph{One-command replication.}
For this repository, a single command is provided:
\begingroup
\scriptsize
\begin{verbatim}
make reproduce
\end{verbatim}
\endgroup
This target runs the artifact-generation scripts used by the main manuscript tables and figures (including constrained transfer, uncertainty, efficiency, adversarial perturbation, and qubit-scaling artifacts).
This one-command path is a core subset pipeline for the primary transfer and mechanism-focused artifacts on this host. The extended script list above remains the full regeneration workflow for all auxiliary analyses.

\paragraph{Checklist links.}
The reproducibility checklist used for this paper is versioned in the repository's docs directory as the reproducibility harness file.
Venue-specific submission checklists are tracked at:
\begin{itemize}
    \item \texttt{docs/NEURIPS\_SUBMISSION\_CHECKLIST.md}
    \item \texttt{docs/ICLR\_SUBMISSION\_CHECKLIST.md}
    \item \texttt{docs/IEEE\_SUBMISSION\_CHECKLIST.md}
\end{itemize}
Code and artifacts are available upon request.

\paragraph{Environment notes.}
The reference environment uses Python with fixed seed sets and deterministic script ordering. Although strict bitwise reproducibility across all hardware is not guaranteed, protocol-level reproducibility is enforced through documented commands, structured outputs, and validation checks.

This reproducibility framing aligns with current deployment governance guidance and standards updates through 2024--2025 \citep{nistairmf2023,nistcsf2024}.